---
<<<<<<< HEAD
title: "PA1_template"
---

###Loading and preprocessing the data

Set up any initial parameters, load the data into 'activity' and format the date column

```{r setup}
library(dplyr)
library(lubridate)
library(chron)
par(mfcol = c(1,1))
activity <- tbl_df(read.csv("activity.csv", na.strings = "NA"))
mutate(activity,date = ymd(date))
head(activity)
```

###What is mean total number of steps taken per day?

Histogram of the total number of steps per day

```{r first_histogram}
activity_sum_steps <- summarize(group_by(activity,date), sum = sum(steps, na.rm = TRUE))
hist(activity_sum_steps$sum, main = "Activity", xlab = "Total steps each day", ylab = "Frequency (number of days with those steps)")
```

Calculate the mean and median of the total steps per day

```{r mean_median}
mean(activity_sum_steps$sum)
median(activity_sum_steps$sum)
```

###What is the average daily activity pattern?

Time-series plot of the 5-min interval and the average steps taken averaged across all days

```{r average_daily_activity}
activity_ave_steps <- summarize(group_by(activity,interval), ave = mean(steps, na.rm = TRUE))
plot(activity_ave_steps$interval, activity_ave_steps$ave, type="l",
     main = "Average Daily Activity", xlab = "5-min Interval", ylab = "Average steps taken")
```

5-min interval that has the maximum number of steps, on average

```{r max_interval}
activity_ave_steps[activity_ave_steps$ave==max(activity_ave_steps$ave),]
```

###Inputing missing values

Total number of missing values

```{r total_missing_values}
colSums(is.na(activity))
```

The strategy for filling in the missing values will be to use the mean for that 5-min interval.

Here is the new data set with the missing data filled in using that strategy.

```{r nas_replaced}
activity_nas_replaced <- mutate(activity, steps = ifelse(is.na(activity$steps),activity_ave_steps$ave,activity$steps))
activity_nas_replaced
```

Histogram of the total number of steps taken each day, along with the mean and median.

For comparison purposes, the two graphs are displayed side by side, which shows that adding the missing NA values did impact the results.
By adding in the new values, the histogram increases the frequency significantly for the first and third bars.

```{r nas_replaced_sum_steps}
activity_nas_replaced_sum_steps <- summarize(group_by(activity_nas_replaced,date), sum = sum(steps, na.rm = TRUE))

par(mfcol = c(1,2))
hist(activity_sum_steps$sum, main = "Activity (Original Plot)\nNAs removed",
     xlab = "Total steps each day", ylab = "Frequency (number of days with those steps)", ylim = c(0,40))
hist(activity_nas_replaced_sum_steps$sum, main = "Activity (New Plot)\nNAs replaced with interval mean",
     xlab = "Total steps each day", ylab = "Frequency (number of days with those steps)", ylim = c(0,40))
```

###Are there differences in activity patterns between weekdays and weekends?

Create new factor with "weekday" and "weekend" levels

```{r nas_replaced_with_weekdays}
activity_nas_replaced_with_weekdays <- mutate(activity_nas_replaced, weekdays = ifelse(is.weekend(date),"weekend","weekday"))
table(activity_nas_replaced_with_weekdays$weekdays)
```

2-panel time series plot for both levels

```{r weekdays_breakdown_for_nas_replaced_with_weekdays}
par(mfcol = c(2,1))
activity_nas_replaced_with_weekdays_ave_steps <- summarize(group_by(activity_nas_replaced_with_weekdays,interval,weekdays),
                                                           ave = mean(steps, na.rm = TRUE))
with(subset(activity_nas_replaced_with_weekdays_ave_steps,weekdays == c("weekend")),
     plot(interval,ave, type ="l",
          main = "weekend", xlab = "Interval", ylab = "Number of steps"))
with(subset(activity_nas_replaced_with_weekdays_ave_steps,weekdays == c("weekday")),
     plot(interval,ave, type ="l",
          main = "weekday", xlab = "Interval", ylab = "Number of steps"))     
```

That's all folks!
=======
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
>>>>>>> 80edf39c3bb508fee88e3394542f967dd3fd3270
